Add Example Title Here

### 2024-03-18 examples/_Template_Example_Complete.qmd
updates

---
title: "Difference Scores: Paired Samples T-test"
author: "Biostatistics Working Group"
theme: cosmo
---

@@

This example assesses growth in a subsample of ABCD participants from Baseline (T0) to the 1-Year follow-up (T1), using weight as a representative metric. The analysis is conducted in two primary steps: 1) a difference score is calculated between baseline and Year_1 weight measurements for each participant; 2) a one-sample t-test is used to test whether the difference score is statistically different than a null hypothesis of zero change. Finally, a visual inspection is further conducted via a scatterplot to graphically represent the relationship between participant's weight at Baseline and Year_1. The ensuing analysis and interpretations are detailed in the subsequent sections.

@@

::: blue
> **This code installs the r packages necessary for this example, if
> they are not already installed**

@@

# Create a list of required packages
packages_required <- c("tidyverse", "easystats", "lme4", "gtsummary", "report", "broom", "gridExtra")

@@

library(tidyverse) # Collection of R packages for data science
library(easystats) # Collection of R packages statistical modeling, visualization, and reporting
library(gtsummary) # Publication-ready tables
library(report) # Easy reporting of regression analyses
library(broom) # Tidy and augment statistical models output
library(gridExtra) # Arrange multiple grid-based plots on a page

@@

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE
#| cache: FALSE

# Set the data paths
data_path_1 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/demo5.0.rds"
data_path_2 <- "/Users/shawes/ABCD/data/rds/abcd_5.0_rds/core-rds-5.0/non-imaging_excluding_nt_5.0.rds"

# Read the data
data_demographics <- readRDS(data_path_1)
data_nonimaging <- readRDS(data_path_2)

# Subset the nonimaging data to include desired variables
selected_vars <- c("src_subject_id", "eventname", "nihtbx_totalcomp_fc", "anthroweightcalc", "anthroweightcalc")
subset_data <- data_nonimaging[, selected_vars]

library(dplyr)
# # Merge the datasets on 'src_subject_id' and 'eventname'
merged_data <- data_demographics %>%
    full_join(subset_data, by = c("src_subject_id", "eventname"))

# Inspect the merged data structure
str(merged_data)

# Define event names to be retained in the analysis and convert variables to appropriate data types
eventnames_to_include <- c(
    "baseline_year_1_arm_1",
    "1_year_follow_up_y_arm_1",
    "2_year_follow_up_y_arm_1",
    "3_year_follow_up_y_arm_1",
    "4_year_follow_up_y_arm_1"
)

df <- merged_data %>%
    filter(eventname %in% eventnames_to_include) %>%
    mutate(
        src_subject_id = as.factor(src_subject_id),
        eventname = factor(eventname, levels = eventnames_to_include, ordered = TRUE),
        age = as.numeric(age),
        sex = as.factor(sex),
        race.4level = as.factor(race.4level),
        hisp = as.factor(hisp),
        high.educ.bl = as.factor(high.educ.bl),
        household.income.bl = as.factor(household.income.bl),
        acs_raked_propensity_score = as.numeric(acs_raked_propensity_score),
        rel_family_id.bl = as.factor(rel_family_id.bl),
        site_id_l = as.factor(site_id_l),
        nihtbx_totalcomp_fc = as.numeric(nihtbx_totalcomp_fc),
        anthroweightcalc = as.numeric(anthroweightcalc),
        anthroweightcalc = as.numeric(anthroweightcalc)
    ) %>%
    # Exclude cases from unused assessment waves
    filter(!is.na(eventname))

@@

descriptives_table <- df %>%
    select(eventname, sex, race.4level, hisp, anthroweightcalc) %>%
    mutate(eventname = factor(eventname, labels = c("Baseline", "Year 1", "Year 2", "Year 3", "Year 4"))) %>%
    mutate(sex = factor(sex, labels = c("Female", "Male"))) %>%
    tbl_summary(
        by = eventname,
        missing = "no",
        label = list(
            sex ~ "Sex", race.4level ~ "Race", hisp ~ "Hispanic",
            anthroweightcalc ~ "Weight"
        ),
        statistic = list(all_continuous() ~ "{mean} ({sd}) )", all_categorical() ~ "{p}%"),
    ) %>%
    modify_header(all_stat_cols() ~ "**{level}**<br>N = {n}") %>%
    bold_labels() %>%
    italicize_levels() %>%
    modify_spanning_header(all_stat_cols() ~ "**Assessment Wave**")
theme_gtsummary_compact()

descriptives_table
```

@@

### Results

::: blue
The code snippet below tells R to compute a difference score by subtracting each participant's weight at T1 from their weight at T0. Relevant summary statistics are also provided.

**STEP 1: Compute Difference Score**

```{r build-model-ttest-1}
#| message: FALSE
#| warning: FALSE
#| echo: TRUE
#| class.source: 'fold-hide'

# Function to compute difference scores
compute_difference <- function(df, variable_name) {
    # Define the event names of interest
    baseline_event <- "baseline_year_1_arm_1"
    followup_event <- "1_year_follow_up_y_arm_1"

    # Compute the difference between Baseline and Year 1 data for the given variable
    diff_data <- df %>%
        filter(eventname %in% c(baseline_event, followup_event)) %>% # Filter for specific event names
        select(src_subject_id, eventname, all_of(variable_name)) %>% # Select required columns
        spread(eventname, variable_name) %>% # Convert data from long to wide format
        mutate(diff = get(followup_event) - get(baseline_event)) %>% # Compute difference between the two time points
        mutate(group_resid = (diff) - mean(diff)) %>% # compute residuals (group_resid).
        drop_na(diff) # Exclude rows with NA in the computed difference

    # Summarize the computed difference scores
    diff_summary <- summary(diff_data$diff)

    # Return the difference data and its summary
    list(data = diff_data, summary = diff_summary)
}

# List of variables for which difference scores are to be computed
variables_of_interest <- c("anthroweightcalc")

# Compute and store difference scores and summaries for each variable in a list
difference_and_summary_list <- lapply(variables_of_interest, function(var) {
    compute_difference(df, var)
})

# Extract the difference data for the 'anthroweightcalc' variable
weight_diff_data <- difference_and_summary_list[[1]]$data

# Merge the 'diff' column back to the main df using 'src_subject_id' as the key
df <- left_join(df, weight_diff_data %>% select(src_subject_id, diff), by = "src_subject_id")

@@

>>>>

---
title: "Add Example Title Here"
author: "Name Here"
---

@@

[ Insert a brief description of the analysis and its purpose here. ]

@@

::: blue
> **This code installs the r packages necessary for this example, if they are not already installed**

@@

# Create a list of required packages
packages_required <- c("addpackage", "addpackage", "addpackage")

@@

library(addpackage) # snippet about the package
library(addpackage) # snippet about the package
library(addpackage) # snippet about the package

@@

```{r read-data}
#| echo: TRUE
#| messages: FALSE
#| warning: FALSE
#| output: FALSE

# Read and wrangle data

@@

@@

## Results

::: blue
[Brief explanation of the code]
```{r build-model-ttest-1}
#| message: FALSE
#| warning: FALSE
#| echo: TRUE

@@

### Model Summary {.tabset .tabset-fade .tabset-pills}

A plot to show xxxxx.
```{r model-checking}

[brief write-up/explanation of the results]

### Model Plots  {.tabset .tabset-fade .tabset-pills}

::: blue
```{r scatterplot}

## Wrapping Up

::: panel-tabset
### Write-up {.tabset .tabset-fade .tabset-pills}

[insert brief write-up here]
